### 2020-02-21 index.Rmd

---
title: 'Fluent statistical computing interfaces for biological data analysis'
author: 'Stuart Lee'
degreetype: 'Doctor of Philosophy'
site: bookdown::bookdown_site
degrees: 'B.MatComSci., University of Adelaide; MSc.Bioinformatics. University of Melbourne'
bibliography: [bib/plyranges.bib, bib/fluentGenomics.bib, bib/thesisrefs.bib]
link-citations: yes
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)

opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = here::here('img/'), fig.align = 'center', fig.show = 'hold',
  cache = TRUE, external = TRUE, comment = "#>", dpi = 300,
  out.width = ifelse(is_html_output(), "100%", "\\textwidth"),
  cache.path = ifelse(is_html_output(), 
                      here::here('cache-gitbook/'), 
                      here::here('cache-pdfbook/')
  )
)

```

`r if (knitr::is_latex_output()) '<!--'`

# Welcome {-}

This is the website for my PhD thesis at Monash University (Australia), titled "Fluent statistical computing interfaces for biological data analysis". 
It is a work in progress.

`r if (knitr::is_latex_output()) '-->'`