### 2020-07-08 Rmd/06-conclusion.Rmd
first pass at intro to conculsion

# Conclusion  {#ch:conclusion}

In this thesis, I have designed tools to explore workflow steps
that are integral to modern biological data science. In particular, I have
implemented software that facilitates the wrangling, integration, and 
visualisation of high-throughput biological data in a principled and pragmatic
manner. Chapter \@ref(ch:plyranges), "plyranges: a grammar of genomic data
transformation" developed a new domain specific language for genomics data.
Chapter \@ref(ch:fluentGenomics), "Fluent genomics with plyranges and tximeta"
described techniques integrating data along the genome, and emphasised
the importance of interoperability between tools. Chapter \@ref(ch:intron),
"Exploratory coverage analysis with superintronic and plyranges", tackled data 
integration from a different angle by looking at multiple summaries
of variables measured along the genome to find regions of biological interest. 
Chapter \@ref(ch:tsne), "Casting multiple shadows: high-dimensional 
interactive data visualisation with tours and embeddings", explored pragmatic
approaches to high dimensional data visualisation in light of the rise of 
popular non-linear embedding methods.

The chapters

<!-- remind the reader of the research problem and purpose and how they were addressed -->
<!-- briefly summarise what has been covered in the paper -->
<!-- make some kind of holistic assessment/judgement/ claim that pertains to the whole project (i.e., more than a descriptive summary) -->
<!-- assess the value/relevance/ implications of the key findings in light of existing studies and literature -->
<!-- ‘speak’ to the Introduction -->
<!-- outline implications of the study (for theory, practice, further research) -->
<!-- comment on the findings that failed to support or only partially support the hypothesis or research questions directing the study -->
<!-- refer to the limitations of the studies that may affect the validity or the generalisability of results -->
<!-- make recommendations for further research -->
<!-- make claims for new knowledge/ contribution to knowledge. -->

## Findings and limitations

## Software development and impact

## Further research