### 2020-07-17 Rmd/00-b-abstract.Rmd
first pass at abstract

# Abstract {-}

Exploratory data analysis is vital to modern science workflows; it allows
scientists to grasp problems with their data and generate new hypotheses.
This work explores three facets of exploratory data analysis workflows as 
applied to biological data science: data wrangling, integration and 
visualisation.  It contributes new statistical computing interfaces and 
frameworks with the explicit aim of enabling scientists to understand
their data and models in their biological context. In chapter 
\@ref(ch:plyranges) we show that genomics data can be represented using
tidy data semantics, and consequently the process of wrangling it can
be simplified via our grammar of genomic data transformation. 
The next contribution is exploring the implications
of our grammar on the integration and representation of genomics data.
In chapter \@ref(ch:fluentGenomics), we provide a framework for integrating
genomics data from multiple assays, via combining model estimates over
their genomic regions. Next we extend our grammar to represent single
variable measurements along the genome in multiple ways; 
in chapter \@ref(ch:intron) we present a software tool that allows coverage
scores to be aggregated and visualised over an experimental design and 
genomic features and use this tool to uncover intron signal in RNA-seq data.
Finally, in chapter \@ref(ch:tsne) we contribute a new visualisation 
interface that provides scientists with a toolkit for discovering structure in 
their high dimensional data, and assist them in understanding when non-linear 
dimension reduction has worked appropriately.

\newpage