### 2020-07-29 Rmd/05-embeddings.Rmd
update chapter 5

```{r setup-videos, cache = FALSE}
vid_lookup <- tibble::tribble(
  ~key, ~url, ~path,
  "gaussian",  "https://player.vimeo.com/video/439635921", here::here("img", "liminal-screenshot-gaussian.png"),
  "hierarchical", "https://player.vimeo.com/video/439635905", here::here("img", "liminal-screenshot-hierarchical.png"),
  "trees-01", "https://player.vimeo.com/video/439635892", here::here("img", "liminal-screenshot-trees-01.png"),
  "trees-02", "https://player.vimeo.com/video/439635863", here::here("img", "liminal-screenshot-trees-02.png"),
  "mouse-01", "https://player.vimeo.com/video/439635812", here::here("img", "liminal-screenshot-mouse-01.png"),
  "mouse-02", "https://player.vimeo.com/video/439635843", here::here("img", "liminal-screenshot-mouse-02.png")
)

display_vid_html <- function(key, lookup) {
  video <- lookup[lookup$key == key, ]
  knitr::include_url(video[["url"]])
}

display_screenshot <- function(key, lookup) {
  img <- lookup[lookup$key == key, ]
  knitr::include_graphics(img[["path"]])
}

```

# Casting multiple shadows: high-dimensional interactive data visualisation with tours and embeddings {#ch:tsne}

There has been a rapid uptake in the use of non-linear dimensionality reduction
(NLDR) methods such as t-distributed stochastic neighbour embedding (t-SNE) in
the natural sciences as part of cluster orientation and dimension reduction
workflows. The appropriate use of these methods is made difficult by their
complex parameterisations and the multitude of decisions required to balance
the preservation of local and global structure in the resulting visualisation.
We present visual diagnostics for the pragmatic usage of NLDR methods
by combining them with a technique called the tour. A tour is a sequence of
interpolated linear projections of multivariate data onto a lower dimensional
space. The sequence is displayed as a dynamic visualisation, allowing a user to
see the shadows the high-dimensional data casts in a lower dimensional view. By
linking the tour to a view obtained from an NLDR method, we can preserve global
structure and through user interactions like linked brushing observe
where the NLDR view may be misleading. We display several case studies from
both simulated and real data from single cell transcriptomics, that shows our
approach is useful for cluster orientation tasks.  The implementation of our 
framework is available as an R package called **liminal** available at
https://github.com/sa-lee/liminal.

## Introduction

High dimensional data is increasingly prevalent in the natural sciences and
beyond but presents a challenge to the analyst in terms of data cleaning,
pre-processing and visualisation. Methods to embed data from a high-dimensional
space into a low-dimensional one now form a core step of the data analysis
workflow where they are used to ascertain hidden structure and de-noise data
for downstream analysis .

Choosing an appropriate embedding presents a challenge to the analyst. How does
an analyst know whether the embedding has captured the underlying topology and
geometry of the high dimensional space? The answer depends on the analyst's
workflow. @Brehmer2014-hk characterised two main workflow steps that an
analyst performs when using embedding techniques: dimension reduction and
cluster orientation. The first relates to dimension reduction achieved by using
an embedding method, here an analyst wants to characterise and map meaning onto
the embedded form, for example identifying batch effects from a high throughput
sequencing experiment, or identifying a gradient or trajectory along the
embedded form [@Nguyen2019-yh]. The second relates to using embeddings as part of
a clustering workflow. Here analysts are interested in identifying and naming
clusters and verifying them by either applying known labels or colouring by
variables that are a-priori known to distinguish clusters. Both of these
workflow steps rely on the embedding being representative of the original high
dimensional dataset, and becomes much more difficult when there is no underlying
ground truth.

As part of a visualization workflow, it's important to consider the perception
and interpretation of embedding methods as well. @Sedlmair2013-pn showed that
2D scatter plots were mostly sufficient for detecting class separation, however
they also noted that often multiple embeddings were required. For the task of
cluster identification, @Lewis2012-ai showed experimentally that novice users
of non-linear embedding techniques were more likely to consider clusters of
points on a 2D scatter plot to be the result of a spurious embedding compared
to advanced users who were aware of the inner workings of the embedding
algorithm.

A complementary approach for visualizing structure in high dimensional data is 
the tour. A tour is a sequence of projections of a high dimensional dataset
onto a low-dimensional basis matrix, and is represented as a
animated visualization [@Asimov1985-vp; @Buja1986-zr]. Given the dynamic nature
of the tour, user interaction is  important for controlling and exploring the 
visualisation: the tour has been used previously by @Wickham2015-cx and  
exploring statistical model fits and by @Buja1996-fk for exploring the space of 
factorial experimental designs.

While there has been much work on the algorithmic details of
embedding methods, there are relatively few tools designed to assist users
to interact with these techniques: when is an embedding sufficient for the
task at hand? Several
interactive interfaces have been proposed for evaluating or using embedding
techniques. @Buja2008-fn used tours to guide analysts during the optimisation
of multidimensional scaling methods by extending their interactive visualisation
software called **XGobi** and **GGobi** into a new tool called **GGvis** 
[@Swayne1998-uq; @Swayne2003-qd; @Swayne2004-ua].
Their interface allows the analyst to dynamically modify and check whether
an MDS configuration has preserved the locality and closeness
of points between the configuration and the original data.
@Ovchinnikova2020-sy created the **Sleepwalk**  interface for checking
non-linear embeddings in single cell RNA-seq data. It provides a click and 
highlight visualisation for colouring points in an embedding according to an 
estimated pairwise distance in the original high-dimensional space.
Similarly, the **TensorFlow**  embedding projector is a web interface to 
running some non-liner embedding methods live in the browser and 
provides interactions to colour points, and select nearest neighbours
[@Smilkov2016-hp]. Finally, the work by @Pezzotti2017-cz provides a user 
guided and modified form of the  t-SNE algorithm, that allows users to modify 
optimisation parameters in real-time.

There is no one-size fits all: finding an appropriate embedding for a given
dataset is a difficult and a somewhat poorly defined problem. For non-linear
methods, there are a lot of parameters to explore that can have an effect on
the resulting visualisation and interpretation. Interfaces for evaluating
embeddings require interaction but should also be able to be incorporated into
an analysts workflow. Instead, we implement a more pragmatic workflow by
incorporating interactive graphics and tours with embeddings that allows users
to see a global overview of their high dimensional data and assists them with
cluster orientation tasks.

The rest of the paper is organised as follows. The next section provides
background on dimension reduction methods, including an overview of the tour.
Then we describe the visual design of **liminal**, followed by implementation
details. Next we provide case studies that show how our interface assists
in using embedding methods. Finally, we describe the insights gained by using
**liminal** and plans for extensions to the software.

## Overview of Dimension Reduction

In the following section, we provide a brief overview of the goals of DR
methods with respect to the data analyst, in addition to their high level the
mathematical details. Here we restrict our attention to two recent methods that
are commonly used in the literature: t-distributed stochastic neighbour embedding 
(t-SNE) and uniform manifold alignment and projection (UMAP); however we do 
mention several other techniques [@Maaten2008-sk; @McInnes2018-co].

To begin we suppose the data is in the form rectangular
matrix of real numbers, $X = [\mathbf{x_1}, \dots, \mathbf{x_n}]$, where $n$ is
the number of observations in $p$ dimensions.  The purpose of any DR algorithm
is to find a low-dimensional representation of the data, 
$Y = [\mathbf{y_1}, \dots, \mathbf{y_n}]$, such that $Y$ is an $n \times d$
matrix where $d \ll p$. The hope of the analyst is that the DR procedure to 
produces $Y$ will remove noise in the original dataset while retaining any 
latent structure.

DR methods can be classified into two broad classes: linear and non-linear
methods. Linear methods perform a linear transformation of the data, that is,
$Y$ is a linear transformation of $X$; one example is principal components
analysis (PCA) which performs an eigendecomposition of the estimated sample
covariance matrix. The eigenvalues are sorted in decreasing order and represent
the variance explained by each component (eigenvector). 
A common approach to deciding on the number of principal components to retain is
to plot the proportion of variance explained by each component and choose a 
cut-off.

For non-linear methods $Y$ is generated via a pre-processed form of the input
$X$ such as the $k$-nearest neighbours graph or via a kernel transformation.
Multidimensional scaling (MDS) is a class of DR methods that aims to construct
an embedding  $Y$ such that the pair-wise distances (inner products) in $Y$ 
approximate the pair-wise distances (inner products) in $X$ 
[@Torgerson1952-am; @Kruskal1964-cz]. There are many 
variants of MDS, such as non-metric scaling which amounts to replacing
distances with ranks instead [@Kruskal1964-cw]. 
A related technique is Isomap which uses a $k$-nearest neighbour graph
to estimate the pair-wise geodesic distance of points in $X$ then uses classical 
MDS to construct $Y$ [@Silva2003-xw]. Other approaches are based on diffusion
processes such as diffusion maps [@Coifman2005-ak]. A recent example of this 
approach is the PHATE algorithm [@Moon2019-ce]. Here an affinity matrix is 
estimated via the pair-wise distance matrix and k-nearest neighbours graph of $X$. 
The algorithm de-noises estimated distances in high-dimensional space via 
transforming the affinity matrix into a Markov transition probability matrix 
and diffusing this matrix over a fixed number of time steps. Then the diffused 
probabilities are  transformed once more to construct a distance matrix, 
and classical MDS is  used to generate $Y$. A general difficulty with using 
non-linear DR methods for exploratory data analysis is selecting and tuning 
appropriate parameters. To see the extent of these choices 
we now examine the underpinnings of t-SNE and UMAP.

The t-SNE algorithm estimates
the pair-wise similarity of (Euclidean) distances of points in a high
dimensional space using a Gaussian distribution and then estimates a
configuration in the low dimensional embedding space by modelling similarities
using a t-distribution with 1 degree of freedom [@Maaten2008-sk]. There are several subtleties
to the to use of the algorithm that are revealed by stepping through its
machinery.

To begin, t-SNE transforms pair-wise distances between $\mathbf{x_i}$ and
$\mathbf{x_j}$ to similarities using a Gaussian kernel:

$$ p_{i|j} = \frac{\exp(-\lVert \mathbf{x_i - x_j} \rVert ^ 2 /
2\sigma_i^2)}{\sum_{k \ne i}\exp(-\lVert \mathbf{x_j - x_k} \rVert ^ 2 /
2\sigma_i^2)} $$

The conditional probabilities are then normalised and symmetrised to form a
joint probability distribution via averaging:

$$ p_{ij} = \frac{p_{i|j} + p_{j|i}}{2n} $$

The variance parameter of the Gaussian kernel is controlled by the analyst
using a fixed value of perplexity for all observations:

$$ \text{perplexity}_i = \exp(-\log(2) \sum_{i \ne j}p_{j|i}\log_2(p_{j|i})) $$
As the perplexity increases, $\sigma^2_{i}$ increases, until its bounded above
by the number of observations , $n-1$, in the data, corresponding to
$\sigma^2_{i} \rightarrow \infty$. This essentially turns $t-SNE$ into a
nearest neighbours algorithm, $p_{i|j}$ will be close to zero for all
observations that are not in the $\mathcal{O}(\text{perplexity}_i)$
neighbourhood graph of the $i$th observation [@Van_Der_Maaten2014-zn].

Next, the target low-dimensional space, $Y$, pair-wise distances between
$\mathbf{y_i}$ and $\mathbf{y_j}$ are modelled as a symmetric probability
distribution using a Cauchy kernel:

$$ q_{ij} = \frac{w_{ij}}{Z} \\ \text{where } w_{ij} = \frac{1}{ 1 + \lVert
\mathbf{y_i - y_j} \rVert ^ 2} \text{ and } Z = \sum_{k \ne l} w_{kl}. $$

The resulting embedding $Y$ is the one that minimizes the Kullback-Leibler
divergence between the probability distributions formed via similarities of
observations in $X$, $\mathcal{P}$ and similarities of observations in $Y$,
$\mathcal{Q}$:

$$ \mathcal{L(\mathcal{P}, \mathcal{Q})} = \sum_{i \ne j} p_{ij} \log
\frac{p_{ij}}{q_{ij}}$$

Re-writing the loss function in terms of attractive (right) and repulsive
(left) forces we obtain:

$$ \mathcal{L(\mathcal{P}, \mathcal{Q})} = -\sum_{i \ne j} p_{ij}\log w_{ij} +
\log\sum_{i \ne j} w_{ij} $$

So when the the loss function is minimised this corresponds to large attractive
forces, that is, the pair-wise distances in $Y$ are small when there are
non-zero $p_{ij}$, i.e. $x_i$ and $x_j$ are close together. The repulsive force
should also be small when the loss function is minimised, that is, when
pair-wise distances in $Y$ are large regardless of the magnitude of the
corresponding distances in $X$.

Taken together, these details reveal the sheer number of decisions that an
analyst must make. How does one choose the perplexity? How should the
parameters that control the optimisation of the loss function (done with
stochastic gradient descent), like the number of iterations, or early
exaggeration ( a multiplier of the attractive force at the beginning of the
optimisation), or the learning rate be selected?  It is a known problem that
t-SNE can have trouble recovering topology and that configurations can be
highly dependent on how the algorithm is initialised and parameterized
[@Wattenberg2016-ji; @Kobak2019-lm; @Melville2020]. If the goal is cluster
orientation a recent theoretical contribution by @Linderman2019-dq proved that
t-SNE can recover spherical and well separated cluster shapes, and proposed new
approaches for tuning the optimisation parameters. However, the cluster sizes
and their relative orientation from a $t-SNE$ view can be misleading
perceptually, due to the algorithms emphasis on locality.

Another recent method, UMAP, has seen a large rise in popularity (at least in
single cell transcriptomics) [@McInnes2018-co]. It is a method that is related to
LargeVis [@Tang2016-oz], and like t-SNE acts on the k-nearest neighbour graph.
Its main differences are that it uses a different cost function (cross entropy)
which is optimized using stochastic gradient descent and defines a different
kernel for similarities in the low dimensional space. Due to its computational
speed it is possible to generate UMAP embeddings in more than three dimensions.
It appears to suffer from the same perceptual issues as t-SNE, however it supposedly preserves global structure better than t-SNE [@Coenen2019-to].

### Tours explore the subspace of $d$-dimensional projections

The tour is a visualisation technique that is grounded in 
mathematical theory, and is able to ascertain 
the shape and global structure of a dataset via inspection of the subspace generated by the set of low-dimensional projections [@Asimov1985-vp; @Buja1986-zr].

Like, when using other DR techniques, the tour assumes we have a real data matrix
$X$ consisting of $n$ observations in $p$ dimensions. First, the tour generates 
a  sequence of $p \times d$ orthonormal projection matrices (bases) 
$A_{t \in \mathbb{N}}$, where $d$ is typically $1$ or $2$. For each pair of 
orthonormal bases $A_{t}$ and $A_{t+1}$ that are generated, the geodesic
path between them is interpolated to form intermediate frames, and giving the 
sense of continuous movement from one basis to another. The tour is then the 
continuous visualisation of the projections $Y_{t} = XA_{t}$, that is the 
projection of $X$ onto $A_{t}$ as the tour path is interpolated between 
successive bases. A *grand tour* corresponds to choosing new orthonormal
bases at random; allowing a user to ascertain structure via exploring the
subspace of $d$-dimensional projections.  In practice, we first sphere 
our data via principal components to reduce dimensionality of $X$ prior to 
running the tour. Instead of picking projections at 
random, a *guided tour* can be used to generate a sequence of 'interesting' 
projections as quantified by an index function [@Cook1995-bi]. While our software, 
**liminal** is able to  visualise guided tours, our focus in the case studies 
uses the grand tour to see global structure in the data.

## Visual Design

Tours provide a supportive visualisation to NLDR graphics, and 
can be easily incorporated into an analysts workflow with our software package, 
**liminal**.  Our interface allows analysts to quickly compare 
views from embedding methods and see how an embedding method 
preserves or alters the geometry of their data. Using multiple concatenated and
linked views with the tour enhances interaction techniques, and allows 
analysts to perform  cluster orientation tasks via linked highlighting and 
brushing [@McDonald1982-ew; @Becker1987-gd]. 
This approach allows our interface to achieve 
the three principles for interactive high-dimensional data visualisation outlined
by @Buja1996-fk: finding gestalt, posing queries, and making comparisons.

### Finding Gestalt: focus and context

To investigate latent structure and the shape of a high dimensional dataset, a
tour can be run without the use of an external embedding. It is often useful to
first run principal components on the input as an initial dimension reduction
step, and then tour a subset of those components instead, i.e. by selecting
them from a scree plot. The default tour layout is a scatter plot with an axis
layout displaying the magnitude and direction of each basis vector. Since the
tour is dynamic, it is often useful to be able to pause and highlight a
particular view. In addition to pause, play and reset buttons, brushing will 
pause the tour path, allowing users to identify 'interesting' projections. The
domain of  the axis scales from running a tour is called the half range, and is 
computed by rescaling the input data onto hyper-dimensional unit cube. We bind 
the half range to a mouse wheel event, allowing a user to pan and zoom on the 
tour view dynamically. This is useful for peeling back dense clumps of points 
to reveal structure.

### Posing Queries: multiple views, many contexts

We have combined the tour view in a side by side layout with a scatter plot
view as has been done in previous tour interfaces **XGobi** and **DataViewer** 
[@Buja1986-ku; @Swayne1998-uq]. These views are
linked; analysts can brush regions or highlight collections of points in either
view. Linked highlighting can be performed when points have been previously
labelled according to some discrete structure, i.e. cluster labels are
available. This is achieved via the analyst clicking on groups in the legend,
which causes unselected groupings to have their points become less opaque.
Consequently, simple linked highlighting can alleviate a known downfall of
methods such as UMAP or t-SNE: that is distances between clusters are
misleading. By highlighting corresponding clusters in the tour view, the
analyst can see the relationship between clusters, and therefore obtain a more
accurate representation of the topology of their data.

Simple linked brushing is achieved via mouse-click and drag movements. By
default, when brushing occurs in the tour view, the current projection is
paused and corresponding points in the embedding view are highlighted.
Likewise, when brushing occurs in the embedding view, corresponding points in
the tour view are highlighted. In this case, an analyst can use brushing for
manually identifying clusters and verifying cluster locations and shapes:
brushing in the embedding view gives analysts a sense of the shape and
proximity of cluster in high-dimensional space.

### Making comparisons: revising embeddings

As mentioned previously, when using any DR method, we are assuming the
embedding is representative of the high-dimensional dataset it was computed
from. Defining what it means for embedding to be 'representative` or 'faithful'
to high-dimensional data is ill-posed and depends on the underlying task an
analyst is trying to achieve. At the very minimum, we are interested in
distortions and diffusions of the high-dimensional data. Distortions occur when
points that are near each other in the embedding view are far from each other
in the original dataset. This implies that the embedding is not continuous.
Diffusions occur when points are far from each other in the embedding view are
near in the original data.  Whether, points are near or far is reliant on the
distance metric used; distortions and diffusions can be thought of as the
preservation of distances or the nearest neighbours graphs between the
high-dimensional space and the embedding space. As distances can be noisy in
high-dimensions, ranks can be used instead as has been proposed by @Lee2009-zb.
Identifying distortions and diffusions allows an analyst to investigate the
quality of their embedding and revise them iteratively.

These checks are done visually using our side-by-side tour and embedding views.
In the simplest case, a local continuity check can be assessed via one to
one linked brushing from the embedding to the tour view. Similarly, diffusions
are identified from linked brushing on the tour view, highlighting points
in the embedding view.

## Software Infrastructure

We have implemented the above design as an open source R package called
**liminal** [@r-liminal]. The package allows analysts to construct concatenated
visualisations, drawn with the **Vega-Lite** grammar of interactive graphics via
the **vegawidget** package [@Satyanarayan2017-gs; @Lyttle2020-hp]. It provides 
an interface for constructing linked and stand alone interfaces for 
manipulating tour paths via  the **shiny** and **tourr** packages 
[@Chang2020-bq; @Wickham2011-st].

### Tours as a streaming data problem

The process of generating successive bases and interpolating between them to  
construct intermediary frames, means the tour is a dynamic visualisation 
technique. Generally, the user would set $d=2$ the tour is visualised as
a animated scatter plot. This process of constructing bases and intermediate 
frames and visualising the resulting projection is akin to making a "flip book" 
animation. Like with a flip book, an interface to the tour requires the ability
to interact and modify it in real time. The user interface generated in **liminal**
allows a user to play, pause, and reset the tour animation,
panning and zooming to modify the scales of the plot to provide context and
click events to highlight groups of points if a labelling variable has
been placed on the legend.

These interactions are enabled by treating the basis generation as a reactive
stream. Instead of realising the entire sequence, which limits the animation
to have a discrete number of frames, new bases and their intermediate frames
are generated dynamically via pushing the current projection to the 
visualisation interface. The interface listens to events like pressing a button
or mouse-dragging and reacts by pausing the stream. This process allows the
user to manipulate the tour in real time rather than having to fix the number
of bases ahead of time. Additionally, once the user has identified an 
interesting projection or is done with the tour, the interface will return the 
current basis for use downstream.

### Linking and highlighting views via interactions

The embedding and tour views are linked together via rectangular brushes;
when a brush is active, points will be highlighted in the adjacent view. 
Because the tour is dynamic, brush events that become active will pause the
animation, so that a user can interrogate the current view. By default,
brushing on the embedding view will produce a one-to-one linking with the tour
view. For interpreting specific combinations of clusters, the multiple
guides on the legend can be selected in order to see their relative orientations.
The interface is constructed as a **shiny** gadget specifically designed for
interactive data analysis. Selections such as brushing regions and the current
tour path are returned after the user clicks done on the interface and become
available for further investigation.

## Case Studies

The next section steps through case studies cases of our approach using simulated
and real data.

The first three case studies use simulations where the cluster structure and
geometry of the underlying data is known. We start with a simple example
where we generated spherical clusters that are embedded well by t-SNE. Then
we move onto more complex examples where the tour provides insight, such
as clusters that have substructure and where there is more complex geometry
in the data.

In the final case study, we apply our approach to clustering 
the mouse retina data from @Macosko2015-ot, and apply the tour
to the process of verifying marker genes that separate clusters.

The videos presented show the visual 
appearance of the **liminal** interface, and how we can interact with the tour
via the controls previously described. Both code and videos available for 
reproducing this paper are available at https://github.com/sa-lee/paper-liminal.

### Case Study 1: Exploring spherical Gaussian clusters

To begin we look at simulated datasets that reproduce known facts
about the  t-SNE algorithm. Our first data set consists of five spherical 5-$d$
Gaussian clusters embedded in 10-$d$ space, each cluster has the same 
covariance matrix. We then computed a t-SNE layout with default settings 
using the **Rtsne** package [@Rtsne],
and set up the **liminal** linked interface with grand tour on the 10-$d$
observations.

(ref:liminal-gaussian) Screenshots of the **liminal** interface applied to well clustered data, a video of the tour animation is available at https://player.vimeo.com/video/439635921. *(a)*: t-SNE (left) and tour (right) in side by side layout coloured by cluster. *(b)*: brushing the red cluster on the t-SNE view, shows the corresponding points on the tour.

```{r gaussian, fig.cap = "(ref:liminal-gaussian)"}
display_screenshot("gaussian", vid_lookup)
```

```{r gaussian-h, eval = knitr::is_html_output()}
display_vid_html("gaussian", vid_lookup)
```

We see in figure \@ref(fig:gaussian) and linked video that t-SNE has 
correctly split out each cluster and laid them out in a star like formation. 
This is agrees with the tour view,
where once we start the animation, the five clusters begin to appear but
generally points are more concentrated in the projection view compared to 
the t-SNE layout (figure \@ref(fig:gaussian)a). 
This can be confirmed via brushing the t-SNE view (figure \@ref(fig:gaussian)b).

### Case Study 2: Exploring spherical Gaussian clusters with hierarchical structure

We repeat the process of viewing Gaussian clusters from the 
_Multi Challenge Dataset_, a benchmark simulation data set for clustering tasks
[@Rauber2009-vh]. This dataset has two Gaussian clusters with equal covariance 
embedded in $10-d$, and a third cluster
with hierarchical structure. This cluster has two 3-$d$ clusters embedded in 
10-$d$, where the second cluster is subdivided into three smaller clusters, 
that are each equidistant from each other and have the same covariance structure.
Again t-SNE has correctly identified the sub-clusters but their relative 
locations to each other is distorted, with the orange and blue groups being
far from each other in the tour view (figure \@ref(fig:hierarchical)a). We see 
in this case that is difficult
to see the sub-clusters in the tour view, however, once we zoom and highlight
they become more apparent (figure \@ref(fig:hierarchical)b). When we brush 
the sub-clusters in the t-SNE, their
relative placement is again exaggerated, with the tour showing that they
are indeed much closer than the impression the t-SNE view gives.

(ref:liminal-hierarchical) Screenshots of the **liminal** interface applied to sub-clustered data, a video of the tour animation is available at https://player.vimeo.com/video/43963590. *(a)* t-SNE has identified and spread out the clusters with substructure. *(b)* Using linked highlighting and brushing, the tour view shows that sub-clusters are much closer together in high dimensional space than the t-SNE view implies.

```{r hierarchical, fig.cap="(ref:liminal-hierarchical)"}
display_screenshot("hierarchical", vid_lookup)
```

```{r hierarchicall-h, eval = knitr::is_html_output()}
display_vid_html("hierarchical", vid_lookup)
```

### Case Study 3: Exploring data with piecewise linear structure

Next explore some simulated noisy tree structured data (figure
\@ref(fig:fake-trees)), our interest here is how t-SNE visualisations break
topology of the data, and then seeing if we can resolve this by tweaking the
default parameters with reference to the global view of the data set.

(ref:fake-trees) Example high-dimensional tree shaped data, $n = 3000$ and $p = 100$. *(a)* The true data lies 2-d tree like structure consisting of ten branches. This data is available in the **phateR** package and is simulated via diffusion-limited aggregation (a random walk along the branches of the tree) with Gaussian noise added [@Moon2019-ce]. *(b)* The first two principal components, which form the initial projection for the tour, note that the backbone of the tree is obscured by this view. *(c)* The default t-SNE view breaks the global structure of the tree. *(d)* Altering t-SNE using the first two principal components as the starting coordinates for the embedding, results in clustering the tree at its branching points.

```{r fake-trees, echo = FALSE, fig.align = "center", fig.cap="(ref:fake-trees)"}
knitr::include_graphics(here::here("img", "tree-view.png"))
```

First, we apply principal components and restrict the results down to the first
twelve principal components (which makes up approximately 70% of the variance
explained in the data) to use with the grand tour.

Moreover, we run t-SNE using the default arguments on the complete data (this
keeps the first 50 PCs, sets the perplexity to equal 30 and performs random
initialisation). We then create a linked
tour with t-SNE layout with **liminal** as shown in figure \@ref(fig:trees-01).

(ref:trees-01) Screenshots of the **liminal** interface applied to tree structured data, a video of the tour animation is available at https://player.vimeo.com/video/439635892. *(a)*: Default t-SNE has difficulties embedding the tree. *(b)* linked highlighting shows that the backbone of the tree has been broken by the embedding.

```{r trees-01, fig.cap = "(ref:trees-01)"}
display_screenshot("trees-01", vid_lookup)
```

```{r trees01-h, eval = knitr::is_html_output()}
display_vid_html("trees-01", vid_lookup)
```

We see that the t-SNE view has been unable to recapitulate the
topology of the tree - the backbone (blue) branch has been split into three
fragments (figure \@ref(fig:trees-01)a). We can see this immediately via the linked highlighting over both
plots. If we click on the legend for the zero branch, the blue coloured points
on each view are highlighted and the remaining points are made transparent.
From, here it becomes apparent from the tour view that the blue branch forms
the backbone of the tree and is connected to all other branches. We can also
again see that cluster sizes formed via t-SNE can be misleading; from the tour
view there is a lot of noise along the branches, while this does not appear to
be the case for the t-SNE result (figure \@ref(fig:trees-01)b).

From the first view, we modify the inputs to the t-SNE view, to try and produce
a better trade-off between local structure and retain the topology of the data.
We keep every parameter the same except that we initialise $Y$ with the first
two PCs (scaled to have standard deviation 1e-4) instead of the default random
initialisation and increase the perplexity from 30 to 100. We then combine
these results with our tour view as displayed in figure \@ref(fig:trees-02).

(ref:trees-02) Screenshots of the **liminal** interface applied to tree structured data, a video of the tour animation is available at https://player.vimeo.com/video/439635863.

```{r trees-02, fig.cap = "(ref:trees-02)"}
display_screenshot("trees-02", vid_lookup)
```

```{r trees-02-h, eval = knitr::is_html_output()}
display_vid_html("trees-02", vid_lookup)
```

Figure \@(ref:trees-02) shows that this selection of parameters results in the 
tips of the branches (the three black dots in figure \@ref(fig:fake-trees)a) 
being split into three clusters representing the terminal branches of the tree. 
However, there are perceptual issues following the placement of the three 
groupings on the t-SNE view that become apparent via simple linked brushing. 
If we brush the
tips of the yellow and brown branches (which appear to be close to each other
on the t-SNE view), we immediately see the placement is distorted in the t-SNE
view, and in the tour view these tips are at opposite ends of the tree 
(figure \@ref(fig:trees-02)b).
Although, this is a known issue of the t-SNE algorithm, we can easily identify
it via simple interactivity without knowing the inner workings of the method.

### Case Study 4: Clustering single cell RNA-seq data

A common analysis task in single cell studies is performing clustering to
identify groupings  of cells with similar expression profiles. Analysts
in this area generally use non linear DR methods for verification and identification
of clusters and developmental trajectories (i.e., case study 1). For clustering,
workflows the primary task is verify the existence of clusters and then begin to 
identify the clusters as cell types using the expression of "known" 
marker genes. Here a 'faithful' a embedding should ideally preserve the topology
of the data; cells that correspond to a cell type should lie in the same
neighbourhood in high-dimensional space. In this case study we use our 
linked brushing approaches to look at neighbourhood preservation
and look at marker genes through the lens of the tour.

First, we downloaded the raw mouse retinal single cell RNA-seq data from
@Macosko2015-ot using the **scRNAseq** Bioconductor package [@scRNAseq-d]. We
have followed a standard workflow for pre-processing and normalizing this data 
(described by @Amezquita2020-at): we performed QC using the the `scater`
by removing cells with high proportion of mitochondrial gene expression and low
numbers of genes detected, we log-transformed and normalised the expression 
values and  finally selected highly variable genes (HVGs) using `scran` 
[@McCarthy2017; @Lun2016]. The top ten percent of HVGs were used to subset
the normalised expression matrix and compute PCA using the first 25 components.
Using the PCs we built a shared nearest neighbours graph (with $k = 10$) and 
used Louvain clustering to generate clusters [@Blondel2008-bx].

To check and verify the clustering we construct a **liminal** view. 
We tour the first five PCs (approximately 20% of the variance in expression),
along side the t-SNE view which was computed from all 25 PCs. 
Due to latency of the interface we do weighted sample of rows based on
cluster membership, leaving us with approximately 10 per cent of the original
data size - 4,590 cells. Although this is not ideal, it still allows us to
get a sense of the shape of the clusters as seen in figure \@ref(fig:mouse-01).

(ref:mouse-01) Screenshots of the **liminal** interface applied to single cell data. A video of the tour animation is available at https://player.vimeo.com/video/439635812. *(a)*: linked brushing identifies distortions in cluster placement by the embedding. *(b)*: t-SNE spreads out larger clusters like cluster 2, 4, 6, and 11 which are concentrated in the tour view. *(c)*: the tour view gives a sense of the geometry of the data; the PCs appear to form a tetrahedron like shape. *(d)*: brushing on the tour view highlights diffusions, points in cluster 9 are close together in the tour view are far away in the t-SNE view.

```{r mouse-01, fig.cap = "(ref:mouse-01)" }
display_screenshot("mouse-01", vid_lookup)
```

```{r mouse-01-h, eval = knitr::is_html_output()}
display_vid_html("mouse-01", vid_lookup)
```

We observe that the embedding has mostly captured the
clusters relative location to each other to their location in high dimensional space, with a notable exception of points in cluster 3 and 10 as shown with
linked brushing (figure \@ref(fig:mouse-01)a). As expected, t-SNE mitigates the crowding problem that is
an issue for tour in this case, where points in clusters 2,4,6, and 11 are
clumped together in tour view, but are blown up in the embedding view (figure \@ref(fig:mouse-01)b). 
The tour appears to form a tetrahedron like shape, with points lying on
surface and along the vertices of the the tetrahedron in 5-$d$ PCA space -
a phenomena that has also been observed in @Korem2015-af 
(figure \@ref(fig:mouse-01)c). Brushing on
the tour view, reveals points in cluster 9 that are diffuse in the embedding 
view, points in cluster 9 are relatively far away and spread apart from other 
clusters in the tour view, but has points placed in cluster 3 and 9 in the embedding
(figure \@ref(fig:mouse-01)d).

Next, we identify marker genes for clusters using one sided Welch t-tests with
a minimum log fold change of one as recommended by @Amezquita2020-at, which
uses the testing framework from @McCarthy2009-qx. We select the top 10 marker 
genes that are upregulated in cluster 2, which was one of the clumped clusters
when we toured on principal components. Here the tour becomes an alternative
to a standard heatmap view for assessing shared markers; the basis
generation (shown as the biplot on the left view) reflects the relative 
weighting of each gene. We run the tour directly on the 
log normalised expression values using the same subset as before figure \@ref(fig:mouse-02).

(ref:mouse-01) Screenshots of the **liminal** tour applied to a marker gene set. A video of the tour animation is available at https://player.vimeo.com/video/439635843. *(a)*: the axis (left) and tour (right) view of 10 marker genes, using linked highlighting we see that the marker genes separate clusters 2, 4, 6, and 11 from the other groups. *(b)*: after zooming, it appears that the marker genes give rise to a gradient along the clusters.

```{r mouse-02, fig.cap = "(ref:mouse-02)"}
display_screenshot("mouse-02", vid_lookup)
```

```{r mouse-02-h, eval = knitr::is_html_output()}
display_vid_html("mouse-02", vid_lookup)
```

We see that the expression of the marker genes, appear to separate
the previously clumped clusters 2,4,6, and 11 from the other clusters, 
indicating that these genes are expressed in all four clusters (figure \@ref(fig:mouse-02)a). Using zooming, we can see a trajectory forming along
the clusters, while the axis view shows that magnitude of expression in 
the marker genes is similar across these separated clusters which is 
consistent with the results of marker gene analysis (figure \@ref(fig:mouse-02)b).

## Discussion

We have shown that the use of tours as a tool for interacting with
high dimensional data provides an additional insight for interrogating views
generated from embeddings. The interface we have designed in the **liminal**
package, allows a user to gain a deeper understanding of an embedding 
algorithm, and rectifies perceptual issues associated with NLDR methods via 
linked interactions with the tour. As we have shown in the simulation case
studies, the t-SNE method can produce misleading embeddings but can be
detected through the use linked brushing and highlighting. In the case when
the data has a piecewise linear geometry, like the tree simulation, the tour
preserves the shape of the data which can be obscured by the embedding method.

Our holistic framework is also useful in real data situations, as displayed
in our second case study. The tour when combined with t-SNE allowed us
to identify clusters, while giving us an idea of their orientation to 
each other. Moreover, we could visually inspect the separation of clusters
using a tour on marker gene sets. We see our approach as being valuable
to the single cell analyst who wants to make their embeddings more 
interpretable

In future implementations it will be possible to shift from one to one linked
brushing to one to many linked brushing. This form of brushing acts directly on
a $k$-nearest neighbours ($k$-nn) graph computed from a
reference dataset: when a user brushes over a region in the embedding,
all the points that match the graphs edges are selected on the 
corresponding tour view.  The reference data set for computing
nearest neighbours (for example a distance matrix, or the complete data matrix)
can be independent of the tour or embedding views. The estimation of neighbours
can be computed efficiently with the `FNN` or `RcppAnnoy` packages
on the server side [@fnn-pkg; @rcpp-annoy-pkg]. In place of highlighting, one 
could use opacity or binned colour scales to encode distances or ranks instead 
of the neighbouring points. Brush composition, such as 'and', 'or',
or 'not' brushes, could be used to interrogate mismatches between the 
$k$-nn graphs estimated from both the embedding and tour views.

There are some limitations
in using the **liminal** interface for larger datasets. First,
t-SNE avoids the crowding problem, points are separated into distinct regions
on the display canvas, while for the tour, points are concentrated in the
centre of the projection and become difficult to see. 
Second, as $n$ increases the animation becomes slower as there is more data
passing from the server to the client. Both of these issues could be mitigated
through alternate tour paths and displays. The recent work
by @Laa2020-wr is a promising area for further investigation, as well as
work from topological statistics [@Rieck2017-kk; @Genovese2017-iq].

## Acknowledgements {-}

We would like to thank Dr David Frazier, Dr Ursula Laa and Dr Paul Harrison for 
their feedback on this work. The screenshots and images were compiled using the 
**cowplot** and **ggplot2** packages [@r-cowplot; @Wickham2016-gz].