### 2020-07-29 index.Rmd
add lay summary

---
title: 'Fluent statistical computing interfaces for biological data analysis'
author: 'Stuart Lee'
degreetype: 'Doctor of Philosophy'
site: bookdown::bookdown_site
degrees: 'B.MatComSci., University of Adelaide; MSc.Bioinformatics. University of Melbourne'
bibliography: [bib/plyranges.bib, bib/fluentGenomics.bib, bib/superintronic.bib, bib/liminal.bib, bib/thesisrefs.bib]
link-citations: yes
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)

opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = here::here('img/'), fig.align = 'center', fig.show = 'hold',
  external = TRUE, comment = "#>", dpi = 300,
  out.width = ifelse(is_html_output(), "100%", "\\textwidth"),
  cache.path = ifelse(is_html_output(), 
                      here::here('cache-gitbook/'), 
                      here::here('cache-pdfbook/')
  )
)

```

`r if (knitr::is_latex_output()) '<!--'`

# Welcome {-}

This work contributes new computing interfaces for the analysis of data 
collected from technologies that measure changes in biology, like gene 
expression. The first interface is a new tool that unites data measured along
the genome under a common language of transformation. This interface enables
analysts to interpret and combine multiple results from different technologies
in a clear and simple way. The second interface provides a framework for 
identifying interesting regions along the genome. The third interface empowers
analysts to interrogate complex models using visual diagnostics. All 
are provided as open source software packages.

`r if (knitr::is_latex_output()) '-->'`