### 2020-08-04 Rmd/00-e-preface.Rmd
style

# Preface {-}

This thesis has been written using R Markdown with the **bookdown**
package [@Xie2016-oa; @Xie2017-ke]. The **renv** package
has been used to create a reproducible environment to build the thesis from source [@renv]. All materials required to compile the thesis are available at https://github.com/sa-lee/thesis. An online version of the thesis is available to read at https://sa-lee.github.io/thesis.

Chapter \@ref(ch:plyranges) has been published in _Genome Biology_. It
won the ACEMS Business Analytics Prize for Best Paper in 2019. Chapter \@ref(ch:fluentGenomics) has been published in _F1000 Research_.  Chapter
\@ref(ch:intron) is based on my software contributions to the submitted
article @Lee2019-mf. Chapter \@ref(ch:tsne) has not been submitted.

\clearpage\pagenumbering{arabic}\setcounter{page}{0}