### 2020-10-27 response/response.Rmd
starting to work through comments

---
title: Thesis Amendments
author: Stuart Lee
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

First, I would like to thank my examiners for their constructive feedback and insights. I am excited to incorporate their ideas into my future research. The point by point description of changes are below: the examiners' comments are in red and my response is in black.

## Professor Levi Waldron {-}

Thank you for your comments and identification of typographic errors.

> p. 1 Introduction: it’s -> its

Amended.

> p. 16 “Using plyranges and the the”

Amended.

> p. 34: use of `dir` for a directory path string when `dir()` is a function returning contents of a directory path could be confusing. However, this chapter is already peer-reviewed and published so I would not change it.

Noted. I can change this in the workflow for next Bioconductor release.

> P. 43-44: Why is the genome (“hg38”) set using an endomorphic function, but style (“UCSC”) set using a replacement function (`seqlevelsStyle`)?

At the time of writing, `plyranges` does not have any genome restyling functions, hence why the replacement function was used.

> p. 77: Tensorflow “non-liner embedding” p. 77-78” “a lot of” -> many

Amended.

> p. 80: “its bounded above” ->it’s

Amended.

> p. 82: “Like, when, using” -> Like when using “While our software, liminal is able “ -> While our software, liminal, is able

Amended.

>Section 5.3.1 “Finding Gestalt: focus and context” - this is probably a clever metaphor, but I didn’t understand it.

This was a reference to Buja, Cook, and Swayne (1996) paper mentioned in the paragraph above. Here "Gestalt" refers to Gestalt theory in perception and psychology, it refers to identifying patterns in visual forms.

> p. 84: “Whether, points are near, or far” -> remove both commas

Amended.

>Figure 5.2 video https://player.vimeo.com/video/43963590 does not exist. The URL given in Table 5.1 is different and correct.

Amended.

> p. 98: “We have shown in the case studies, that” -> no comma

Amended.

## Assistant Professor Stephanie C. Hicks {-}

Thank you for your thorough review. I am looking forward to incorporating your suggestions into future work.

### Addressing General Comments

#### On Scalability

I would like to comment that I have avoided performance benchmarking for two main reasons. First, I believe it is out of scope for the thesis, since my focus has been on the development of programming and user interfaces. Second, my implementation of **plyranges** is deeply entwined with the underlying core Bioconductor infrastructure, especially the data structures and methods found in the **IRanges** and **GenomicRanges** packages. This would result any benchmarking of **plyranges** turning into benchmarking of those packages which has been done elsewhere.

Regarding out of memory data structures like `DelayedArray`, it is possible to incorporate them as columns to a `GRanges` object and then the full **plyranges** API would be available. Currently, there appear to be no plans in Bioconductor for out of memory `Ranges`, and indeed there are times where it is more appropriate to switch to a rectangular data layout for computation, as is reflected in the workflow described in chapter 3 of the thesis. I believe the development of a scalable out of memory interface for genomic data analysis will require the transformation between data measured along the genome (covered by the **plyranges** API) and the rectangular layouts that enable out of disk computing like `SummarizedExperiment`. This is an exciting research area that I would like to explore in the currently under development **plyexperiment**.

#### On alignment vs. pseudoalignment

I have amended the introduction to include a note about alignment. 
I believe it is out of scope to go into the gritty details of when to use alignment or pseudo-alignment methods to obtain counts for RNA-seq data. I have included results from both approaches to highlight that **plyranges** is agnostic to the method used and is flexible for many types of data.

### Chapter 1 comments

> As you introduced a “biological data science workflow” in this chapter and did a great job of walking the read through the wrangling, combining and visualizing components of the workflow, I was hoping you would touch on tools for communication too. I understand this is outside the scope of the thesis, but I still think it would be nice to at least mention work others have done in this area.

I agree that it would be nice but it is out of scope for the thesis. Communication is extremely broad and I feel that including it here would distract from the focus of my thesis.

> Pg 1: I thought it was strange that the words “figure” in e.g. “figure 1.1” were not capitalized. Would recommend changing to “Figure 1.1” throughout.
> Pg 1: Recommend capitalizing the word “chapter” when referring to specific chapters e.g. “chapter 2”.

TODO

> Pg 2: Figure 1.1 has been modified to include “combine”, but the caption does not reflect what is meant by that. Would recommend modifying caption to reflect importance of combining experimental assays together to gain biological insight.

Amended.

> Pg 3: you introduce “range-based genomics data”, but it is not clear what is meant by that. More explanation would be helpful.

I have clarified what this refers to.

>Pg 4: In Section 1.2, could you add a few more references on existing methods for integration of genomic data and then re-emphasize the point that however existing methods have not thought much about the interoperability between the tidyverse and Bioconductor approaches?

I have expanded this section to reflect these comments.

### Chapter 2 comments

Most of the comments pertain to the use and learning of the **plyranges** package. I believe they are best addressed through the **plyranges** documentation and vignette. I have amended the thesis to include the package vignette as an appendix.

### Chapter 6 comments

> Pg 102: Instead of https://bioconductor.org/packages/release/bioc/html/plyranges.html, you can always use the short version https://bioconductor.org/packages/plyranges which points to the release version.

Amended.