### 2020-10-27 response/response.Rmd
more details for chapter 5

---
title: Thesis Amendments
author: Stuart Lee
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

First, I would like to thank my examiners for their constructive feedback and insights. I am excited to incorporate their ideas into my future research. The point by point description of changes are below: the examiners' comments are in red and my response is in black.

## Professor Levi Waldron {-}

Thank you for your comments and identification of typographic errors.

> p. 1 Introduction: it’s -> its

Amended.

> p. 16 “Using plyranges and the the”

Amended.

> p. 34: use of `dir` for a directory path string when `dir()` is a function returning contents of a directory path could be confusing. However, this chapter is already peer-reviewed and published so I would not change it.

Noted. I can change this in the workflow for next Bioconductor release.

> P. 43-44: Why is the genome (“hg38”) set using an endomorphic function, but style (“UCSC”) set using a replacement function (`seqlevelsStyle`)?

At the time of writing, `plyranges` does not have any genome restyling functions, hence why the replacement function was used.

> p. 77: Tensorflow “non-liner embedding” p. 77-78” “a lot of” -> many

Amended.

> p. 80: “its bounded above” ->it’s

Amended.

> p. 82: “Like, when, using” -> Like when using “While our software, liminal is able “ -> While our software, liminal, is able

Amended.

>Section 5.3.1 “Finding Gestalt: focus and context” - this is probably a clever metaphor, but I didn’t understand it.

This was a reference to Buja, Cook, and Swayne (1996) paper mentioned in the paragraph above. Here "Gestalt" refers to Gestalt theory in perception and psychology, it refers to identifying patterns in visual forms.

> p. 84: “Whether, points are near, or far” -> remove both commas

Amended.

>Figure 5.2 video https://player.vimeo.com/video/43963590 does not exist. The URL given in Table 5.1 is different and correct.

Amended.

> p. 98: “We have shown in the case studies, that” -> no comma

Amended.

## Assistant Professor Stephanie C. Hicks {-}

Thank you for your thorough review. I am looking forward to incorporating your suggestions into future work.

### Addressing General Comments

#### On Scalability

I would like to comment that I have avoided performance benchmarking for two main reasons. First, I believe it is out of scope for the thesis, since my focus has been on the development of programming and user interfaces. Second, my implementation of **plyranges** is deeply entwined with the underlying core Bioconductor infrastructure, especially the data structures and methods found in the **IRanges** and **GenomicRanges** packages. This would result any benchmarking of **plyranges** turning into benchmarking of those packages which has been done elsewhere.

Regarding out of memory data structures like `DelayedArray`, it is possible to incorporate them as columns to a `GRanges` object and then the full **plyranges** API would be available. Currently, there appear to be no plans in Bioconductor for out of memory `Ranges`, and indeed there are times where it is more appropriate to switch to a rectangular data layout for computation, as is reflected in the workflow described in chapter 3 of the thesis. I believe the development of a scalable out of memory interface for genomic data analysis will require the transformation between data measured along the genome (covered by the **plyranges** API) and the rectangular layouts that enable out of disk computing like `SummarizedExperiment`. This is an exciting research area that I would like to explore in the currently under development **plyexperiment**.

#### On alignment vs. pseudoalignment

I have amended the introduction to include a note about alignment. 
I believe it is out of scope to go into the gritty details of when to use alignment or pseudo-alignment methods to obtain counts for RNA-seq data. I have included results from both approaches to highlight that **plyranges** is agnostic to the method used and is flexible for many types of data.

### Chapter 1 comments

> As you introduced a “biological data science workflow” in this chapter and did a great job of walking the read through the wrangling, combining and visualizing components of the workflow, I was hoping you would touch on tools for communication too. I understand this is outside the scope of the thesis, but I still think it would be nice to at least mention work others have done in this area.

I agree that it would be nice but it is out of scope for the thesis. Communication is extremely broad and I feel that including it here would distract from the focus of my thesis.

> Pg 1: I thought it was strange that the words “figure” in e.g. “figure 1.1” were not capitalized. Would recommend changing to “Figure 1.1” throughout.
> Pg 1: Recommend capitalizing the word “chapter” when referring to specific chapters e.g. “chapter 2”.

TODO

> Pg 2: Figure 1.1 has been modified to include “combine”, but the caption does not reflect what is meant by that. Would recommend modifying caption to reflect importance of combining experimental assays together to gain biological insight.

Amended.

> Pg 3: you introduce “range-based genomics data”, but it is not clear what is meant by that. More explanation would be helpful.

I have clarified what this refers to.

>Pg 4: In Section 1.2, could you add a few more references on existing methods for integration of genomic data and then re-emphasize the point that however existing methods have not thought much about the interoperability between the tidyverse and Bioconductor approaches?

I have expanded this section to reflect these comments.

### Chapter 2 comments

Most of the comments pertain to the use and learning of the **plyranges** package. I believe they are best addressed through the **plyranges** documentation and vignette. I have amended the thesis to include the package vignette as an appendix.

### Chapter 5 comments

>  Pg 77: “non-liner” should be “non-linear”

Amended.

>  Pg 78: “produces” should be “produce”

Amended.

> Pg 78: Could you add a reference for PCA?

Amended.
> Pg 80 (also in 81): The way “tSNE” was written in “this essentially turns t-SNE” looks very different
than previously written stylistically. Could this be converted to the way you were originally referring to
it?

Amended.

> Pg 87: “inn” should be “in”

Amended.

> Pg 88: Inside Figure 5.1A and B, could you have liminal label which plot is which i.e. t-SNE vs tour
view. Also, could axis labels be added? Maybe also add this information to the Figure legend? – Pg 89 – same comment

I think this is not obvious in the static figure but is obvious in the video and when using the **liminal** package that the embedding view is always on the left, and the tour view is on the right (it will be the view that is animated). I think the suggestion of titling the views is a good one and will place that into the software in future. Regarding axis labels, I don't think they make sense for the tour since it's really a linear combination of multiple variables (there's an option to switch on a biplot view for the tour). I am also reluctant to include them for t-SNE because they can be misleading for the reasons mentioned in the chapter, however I do think this would be good to have as an option for **liminal** in future.

> Pg 90: I believe there should be a period after “tree structure data (figure 5.3)” instead of a comma

Amended.

> Pg 94: Can you comment on the scRNA-seq example where you only pick first 5 PCs that explained only 20% of the variance. Or more generally, what is the trade-off on the liminal view by picking more or less PCs along with a NLDR method. Do you have recommendations for the user on how many
PCs to pick for and optimal liminal view?

I have amended section 5.5.3 to provide more details about why the first five components were selected, and how to choose variables to tour.

> Pg 94: Can you comment on the limits of the liminal view in terms of number of observations that
one can visualize in a given tour? Have you thought about strategies for making this scale with more observations?
– Ah, I see you touched on this at the end of the discussion of the capture. I think this is a really important point that would be greatly appreciated going forward.
– And it was talked about in Chapter 6 in Future Work. Incorporating this more in the middle of the chapter with tables specifying what is or is not reasonable to plot in terms of the number of observations, would be helpful.

I have amended section 5.5.3 to discuss a strategy in the specific case of cluster analysis. I have also amended the discussion of the chapter with specific work that can mitigate the need to show all of the data. In the case of tasks performed for clustering I don't think it is useful or necessary to show all points in the scatter plot when we are most interested in density and shapes. There is interesting work to be done here in future for appropriate sampling strategies or aggregations that could be done prior to producing a visualisation.

### Chapter 6 comments

> Pg 102: Instead of https://bioconductor.org/packages/release/bioc/html/plyranges.html, you can always use the short version https://bioconductor.org/packages/plyranges which points to the release version.

Amended.