### 2020-10-28 Rmd/06-conclusion.Rmd

```{r bioc-stats, include = FALSE, cache = TRUE}
library(dplyr)
library(ggplot2)
dl_stats <- BiocPkgTools::biocDownloadStats() %>%
  filter(Package == "plyranges")

plyranges_stats <- dl_stats %>% 
  summarise(
    no_ip = sum(Nb_of_distinct_IPs), 
    no_downloads = sum(Nb_of_downloads)
  )

nice <- function(x) prettyNum(x, big.mark = ",", scientific = FALSE)

```

# Conclusion  {#ch:conclusion}

In this thesis, I have designed tools to explore workflow steps
that are integral to modern biological data science. In particular, I have
implemented software that facilitates the wrangling, integration, and 
visualisation of high-throughput biological data in a principled and pragmatic
manner.  The early chapters of this thesis explored the *tidy data* semantic and 
its extension to range based genomics data. This culminated in the development
of "plyranges: a grammar of genomic data transformation" in Chapter 
\@ref(ch:plyranges), which developed a new domain specific language for genomics
data analysis. The applicability of the **plyranges** interface and use of the 
*tidy data* concept were further interrogated in Chapter \@ref(ch:fluentGenomics),
"Fluent genomics with plyranges and tximeta", which described techniques 
integrating data along the genome, and emphasised the importance of 
interoperability between analysis tools. Similarly, Chapter  \@ref(ch:intron), 
"Exploratory coverage analysis with superintronic and 
plyranges", tackled data integration from a different angle by looking at 
multiple summaries of variables measured along the genome to find putative 
regions of intron retention. In the final part of the thesis I moved towards 
visualisation issues as they related to working with high-dimensional data
common in biological data science. Chapter \@ref(ch:tsne), "Casting multiple shadows:  high-dimensional interactive data visualisation with tours and embeddings",  explored pragmatic approaches to high dimensional data 
visualisation in light of the rise of popular non-linear embedding methods.

A significant amount of my work has been devoted to the development of 
open source R packages and workflows: **plyranges**, **fluentGenomics**, **superintronic** and **liminal**. I have emphasised how coherent software packages are tools for thought; they enable analysts to reason about their data and  models through the composition of workflows. To finish, I will discuss the implications of this work and provide suggestions for further research.

## Software Development

The **plyranges** package develops a suite of verbs for interacting
with genomic data as a *GRanges* object. Since its release on Bioconductor,
it  has been relatively  successful: it has been downloaded `r nice(plyranges_stats$no_downloads)` times from  `r nice(plyranges_stats$no_ip)` unique IP addresses. I have also had the privilege of teaching workshops on **plyranges**
at Bioconductor conferences which also led to the development of the 
**fluentGenomics** workflow package, outlined in Chapter \@ref(ch:fluentGenomics). A broader impact of the work, has been the discussions around the
concepts of fluent interfaces and tidy data within the Bioconductor community,
which has led to several developments currently in place that are exploring
different approaches for fluent interfaces for other types of omics data.
The **plyranges** package is available to download from 
https://bioconductor.org/packages/plyranges and
the **fluentGenomics** workflow is available to download from 
https://bioconductor.org/packages/release/workflows/html/fluentGenomics.html.

The **superintronic** software described in Chapter \@ref(ch:intron) has
been used in @Lee2019-mf to disentangle and view intron signal in RNA-seq
data. Here, we again show the strengths of providing a long-form representations
of genomics data (in this case coverage vectors). By leveraging **plyranges**
we were then able to create a set of data descriptors that we could link back
to the raw data to discover genes thought to be associated with a real biological 
signal. An interesting extension to this work would be applying it to single
cell and long-read based transcriptomics data, where scalability and much
larger design matrices would become an issue. The **superintronic** package
is available to download from https://github.com/sa-lee/superintronic.

Finally, the **liminal** software aims to provide a more holistic approach to
analysis tasks requiring the use of dimensionality reduction algorithms. We
showed how to incorporate interactive graphics and tours to identify problems
with embeddings. Based on the case studies provided I believe that the methods
used in **liminal** could be broadly applicable to many high dimensional datasets
and NLDR methods.
The **liminal** package is available to download from https://github.com/sa-lee/liminal.

## Further Work

A limitation of the grammar as we have implemented it in 
**plyranges** is lack of scalability and computational speed for data sets
that do not fit in memory. We attempted several techniques for performing 
delayed operations over range-based data, however a more general approach
that allows for data stored on the cloud or in scientific data formats
like HDF5 that leverage existing Bioconductor frameworks would be useful.
We showed in chapters \@ref(ch:fluentGenomics) ad \@ref(ch:intron) that an
analyst is able to do some very complex data transformations and re-sampling 
procedures via casting results into _GRanges_ object. However, it is unclear 
whether the semantics of our grammar can be extended to data that can not be 
efficiently reshaped into long form tidy representations. Moreover, further work
is required to explore the design space of grammars for data 
transformations and grammars for graphics when the data are large, multifaceted
and non-rectangular.

We showed in Chapter \@ref(ch:tsne) that tours provide a global overview
that can be used as tool for exploring model fits. An issue that arises is
how to scale the tour as the number of observations increases. There are
latencies in sending data from the back end to the visualisation client
that causes lag during animation. One could also question whether point
based displays are appropriate in this case, and it would be worth exploring 
the usability of animations based on binning the projections. Moreover, 
when the number of observations are large, the points in the projections are 
concentrated in the centre of the tour display obscuring interesting aspects
of the data. This is mitigated via having
the ability to zoom, but further research into transforming the projections
to avoid crowding would be valuable. An added complexity to changes in visual
displays are thinking about the design of user interactions, and several promising
avenues based on section tours could be explored [@Laa2020-wr; @Laa2020-fy].